Algebraic Geometry\vspace{-1ex}

### 2019-11-01 Algebraic Geometry/Algebraic-Geometry.tex
update

>>>>

\section{October 30th, 2019}
Let's start by discussing the idea that for any final object $\Omega$, $X\times_\Omega Y\cong X\times Y$. One way to do this 
is to show that $X\times Y$ also has the universal property corresponding to the limit of $X\to \Omega\leftarrow Y$. But we can 
also do things functorially! Notice that we have a map 
\[h_{X\times_\Omega Y}=h_X\times_{h_\Omega}h_Y\xrightarrow{\cong}h_X\times h_Y=h_{X\times Y}\]
where we can show the isomorphism of these things on points.

Let's talk about the fiber product 
\[\Spec A\times_{\Spec C}\Spec B\]
which intuitively should be something like $\Spec A\otimes_C B$. But is this going to be reduced still? Let's do an example. Consider $A=k[t]$ and $C=k[x]$ 
and consider the map $x\mapsto t^2$ as a map $C\to A$. Then let $B=k$ with the map that sends $x\mapsto 0$. Then $A\otimes_C B\cong k[\varepsilon]/\varepsilon^2$.

The takeaway here is that the tensor product is sometimes non-reduced! It ends up that if we take $(A\otimes_C B)_{red}$ is the algebra we want, which is the 
appropriate object \textit{in the category of reduced $k$-algebras.}
\begin{prob}
	Consider the maps $\Spec(-)$ and $\O(-)$, between the categories $\Alg_k^{op}$ and varieties. These form an adjoint pair! Which object is 
	left adjoint? Do left adjoints commute with limits? colimits?
\end{prob}

\subsection{A series of fun exercises}
Here's some work to do!
\begin{thm}
	Fiber products exist in the category of varieties.
\end{thm}
\begin{rmk}
	You should prove this! We will discuss the idea here. The only thing we know how to do is do this with affines! 
	reduce to the affine case (for all three) and glue!
	\begin{itemize}
		\item Begin by showing: Suppose $U\subseteq Z$ is open and $X\times_Z Y$ exists. Then we can look at 
		\begin{center}
			\begin{tikzcd}
				(X\times_ZY)_U\ar[d]\ar[r,hookrightarrow]& X\times_Z Y\ar[d]\\
				U\ar[r,hookrightarrow] & Z
			\end{tikzcd}
		\end{center}
		and we claim that $(X\times_ZY)_U$ representas $h_{X_u}\times_{h_u}h_{Y_u}$. This gives us a way to shrink over the base.
		\item Now suppose $W\subseteq Y$ is open. Then looking at a similar diagram, we want to show that $(X\times_ZY)_W$ represents $h_X\times_{h_Z}h_W$. The same thing can be done with a subset of $X$.
	\end{itemize}
	Now we want to reduce to $Z$ being affine, then reduce to $Y$ is affine, then $X$. Then we are golden.
	For the first, assume it holds when $Z$ is affine and then show it works for arbitrary $Z$: pick a covering $U_i$ of $Z$ by affines. Consider $U_i$ and $U_j$:

	Each $(X_{U_i}\times_{U_i}Y_{U_i})_{U_i\cap U_j}\cong (X_{U_j}\times_{U_j}Y_{U_j})_{U_i\cap U_j}$ via a canonical isomorphism $\phi_{ij}$. This is the crux of it. THINK. :)
\end{rmk}

\section{November 1st, 2019}
Let's look at some examples of things. $\A1\times\A1=\Spec k[x]\otimes k[y]\cong\Spec k[x,y]\cong \A2$. Another way to see this 
is to think about the functors:
\[\Hom(X,\A1\times\A1)\cong\Hom(X,\A1)\times\Hom(X,\A1)=\Hom(k[s,t],\O(X))=\Hom(X,\A2).\]
Similarly you can do something along these lines with $\A n$.

What about $\P1\times\P1$? It's not $\P2$. Why? Because there are curves $C_1$ and $C_2$ in the product that do not intersect.
But Bezout's theorem says that any two curves in $\P2$ intersect!

The problem with solving the above problem is that we really need sheaf theory to solve the above problem. 
We will circle back around to this.

Question: Is the zariski topology on $X\times Y$ the product topology? NO! For instance $\A2$: the product topology is not the cofinite one!

\subsection{Projective spaces}
\begin{defn}
	A variety is \textbf{projective} if there exists a closed immersion $X\hookrightarrow \P n$.
\end{defn}
\begin{prop}
	Products of projective varieties are projective.
\end{prop}
\begin{prob}
	Classify all projective objects in the category of varieties! You can do it. :)
\end{prob}
\begin{rmk}
	Notice that right now we are working in the \textit{Italian} projective space. This is the usual $\A n\setminus\{0\}$.
	Then we say that $Z$ is closed if the closure of its inverse image in $\A n\setminus \{0\}$ is invariant under scaling.
\end{rmk}
Now think about the functions on $\P n$. There is a ring $R\subseteq k(x_0,\dots,x_n)$ called the field of homogeneous rational functions. These are the 
ratio of homogeneous polynomials of the same degree. It's a field baby. Then 
\[\O_{\P n}(U)=\{\phi\in R|\operatorname{poles}(\phi)\subseteq\P n\setminus U\}.\]

\subsection{The Point}
There is a famous map called the Segre embedding. We get a map 
\[\P n\times\P m\hookrightarrow \P {(n+1)(m+1)}\]
where we can just imagine taking $(v_i)$ and $(w_i)$ and map it to $\mathbf{v}\mathbf{w}^T$.

This is just the projectivization of a matrix! It dends up that it identifies the product with a closed subvariety equal to \textbf{the set of matrices of rank 1.}
There was something at the end about $2\times 2$ subdeterminants!
